---
title:"PA1_template"
output:html_document
---
Course: Reproducible Research PA1_template
Author: Priyantha Perera
Date: August 16 2015

```{r}
setwd("C:/Users/Priyantha/Coursera/Reproducible Research/week2")

library(data.table)
library(stats)
library(dplyr)

```

```{r}
data <- read.csv("activity.csv", stringsAsFactors=FALSE)
head(data)
dim(data)
str(data)
unique(data$date)

datanaomit <- na.omit(data)
dim(datanaomit)
head(datanaomit)
dim(datanaomit)
x <- datanaomit[,2]
class(x)
unique(x)
datetime <- strptime(x, format="%m/%d/%Y", tz="EST")

head(datetime)

datanaomit2 <-cbind(datanaomit,datetime)
names(datanaomit2)
A<-unique(datetime)

```

Total number of steps taken per day calculation
 
```{r}
tmp <-NULL

for (i in 1:length(A)){

  stepsum <- sum(subset(datanaomit,datanaomit2$datetime==A[i])[,1])

  if (is.null(tmp)) tmp<-stepsum else tmp<-c(tmp,stepsum)

}



steps <- as.data.frame(tmp)
steps <- data.frame(A,steps)
names(steps)<- c("date", "steps")
steps
```
Histogram of total number of steps taken each day
```{r}
hist(tmp, breaks = 20)
```

Computing the step means

```{r}
tmpcount <-NULL

for (i in 1:length(A)){

  stepcount <- length(subset(datanaomit,datanaomit2$datetime==A[i])[,1])

  if (is.null(tmpcount)) tmpcount<-stepcount else tmpcount<-c(tmpcount,stepcount)

}


stepmean <- tmp/stepcount

plot(stepmean)

```

Median calculation

```{r}
tmpmed <-NULL
for (i in 1:length(A)){

  stepmed <- median(subset(datanaomit,datanaomit2$datetime==A[i])[,1])

  if (is.null(tmpmed)) tmpmed<-stepmed else tmpmed<-c(tmpmed,stepmed)

}

tmpmed

```

Time series plot of 5 -minute interval avg # of steps taken across all days

```{r}

index <- rep(1:stepcount[1],length(A))
indexeddata <- data.frame(index,datanaomit)
head(indexeddata)

avgstepsin5mtinterval <-NULL

for (i in 1:stepcount[1]){

                tmpts <- sum(subset(indexeddata[,2],indexeddata$index==i))/length(unique(x))


  if (is.null(avgstepsin5mtinterval)) avgstepsin5mtinterval<-tmpts else avgstepsin5mtinterval<-c(avgstepsin5mtinterval, tmpts)


}

plot(avgstepsin5mtinterval)

```

Finding the maximum average and which 5 minute interval it occurs

```{r}
maxsteps <- max(avgstepsin5mtinterval)
maxsteps
rowindexmax <- match(maxsteps, avgstepsin5mtinterval)

```

The maximum of the daily average number of steps in 5 minute intervals occur in the interval:  

```{r}
rowindexmax

```

Missing value handling
We are going to substitute missing values with the average number of steps in the daily 5 minute intervals averaged over all days.

Here is the code and results of total numbr of missing values

```{r}
(dim(data)-dim(datanaomit))[1]
```

This is to visually observe what the missing value data looks like

```{r}
datana <-data[is.na(data[,1]),]

dim(datana)
names(datana)
unique(datana$date)
View(datana)

```

By inspection of datana we observe that there are 8 different days with "NA"s. 

```{r}
(dim(datana)/288)[1]
```

This implies that all 8 days had all the observations missing.
We will replicate "tmpts", the average number of steps  in 5 minute interval "missingdays" = 8 times.

```{r}
missingdays<-dim(datana)[1]/288

missingvaluefill <- rep(avgstepsin5mtinterval, missingdays)

```

Now replace the "steps" column of datana by the "missingvaluefill" values

```{r}
datana[,1] <- missingvaluefill
head(datana)
head(datanaomit)
names(datana)
names(datanaomit)
class(datana)
class(datanaomit)


datana[,1] <- missingvaluefill
head(datana)
head(datana)
names(datana)
class(datana)

datanafilled <-rbind(datanaomit,datana)
head(datanafilled)

y <- datanafilled[,2]
class(y)
unique(y)

datetimeall <- strptime(y, format="%m/%d/%Y", tz="EST")

head(datetimeall)


datanafilled2 <-cbind(datanafilled,datetimeall)
names(datanafilled2)
B<-unique(datetimeall)
length(B)

```

Total number of steps taken per day calculation with na filled data

```{r}

tmpnafilled <-NULL


for (i in 1:length(B)){

  stepsum2 <- sum(subset(datanafilled,datanafilled2$datetimeall==B[i])[,1])

  if (is.null(tmpnafilled)) tmpnafilled<-stepsum else tmpnafilled<-c(tmpnafilled,stepsum2)

}



stepsall <- data.frame(B,tmpnafilled)
names(stepsall)<- c("date", "steps")
stepsall
```


Histogram of total number of steps taken each day

```{r}
hist(tmp, breaks = 20)

```

Computing the daily mean of nafilleddata

```{r}
tmpall <-NULL

for (i in 1:length(B)){
  
  stepsum <- sum(subset(datanafilled,datanafilled2$datetimeall==B[i])[,1])
  
  if (is.null(tmpall)) tmpall<-stepsum else tmpall<-c(tmpall,stepsum)
  
}



stepsall <- as.data.frame(tmpall)
stepsall <- data.frame(B,tmpall)
names(stepsall)<- c("date", "steps")
```
Count
```{r}
tmpnafilledcount <-NULL
dim(datanafilled)
names(datanafilled)

for (i in 1:length(B)){
  
  stepcount <- length(subset(datanafilled,datanafilled2$datetimeall==B[i])[,1])
  
  if (is.null(tmpcount)) tmpcount<-stepcount else tmpcount<-c(tmpcount,stepcount)
  
}

```
Calculating the mean and plotting
```{r}

stepmeanall <- tmpall/stepcount


plot(stepmeanall)

```

Median calculation 
```{r}
tmpmedall <-NULL
for (i in 1:length(B)){
  
  stepmedall <- median(subset(datanafilled,datanafilled2$datetimeall==B[i])[,1])
  
  if (is.null(tmpmedall)) tmpmedall<-stepmedall else tmpmedall<-c(tmpmedall,stepmedall)
  
}

tmpmedall

```

Note: I have not plotted a time series plot, hence the na filled 8 days appear at the end of the plot.


Note That the median of the original data remain at 0, while all the na filled days have the same non zero median.

Remarks:
My method of filling in the NA's is not satisfactory. It does not convey any new information regarding the data set. Perhaps randomly assigning the 5 minute interrval values by matching week days might give a better overall picture of the data

Sorry I was unable to complete the weekday weeekend part 

```{r}
weekdaysall <- weekdays(datanafilled[,2])
head(weekdaysall)

```


